---
title: "Ice cream sales with RStanARM"
author: "John Stanton-Geddes"
date: "May 17, 2016"
output: html_document
---

This analysis complements my talk for the BTV Data Scientists on May 18, 2016. The slides are available at [http://slides.com/johnstanton-geddes/rstanarm#/](http://slides.com/johnstanton-geddes/rstanarm#/)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A comparison of Frequentist and Bayesian methods of a simple example of ice cream sales, borrowed from these [blog](http://www.magesblog.com/2015/08/generalised-linear-models-in-r.html) [posts](http://www.magesblog.com/2015/08/visualising-predictive-distribution-of.html).

The data are the units of ice cream sold at different temperatures.

```{r data}
icecream <- data.frame(
  temp=c(11.9, 14.2, 15.2, 16.4, 17.2, 18.1, 
         18.5, 19.4, 22.1, 22.6, 23.4, 25.1),
  units=c(185L, 215L, 332L, 325L, 408L, 421L, 
          406L, 412L, 522L, 445L, 544L, 614L)
  )
```


```{r basic_plot, echo=FALSE}
basicPlot <- function(...){
  plot(units ~ temp, data=icecream, bty="n", lwd=2,
       main="Number of ice creams sold", col="#00526D", 
       xlab="Temperature (Celsius)", 
       ylab="Units sold", ...)
  axis(side = 1, col="grey")
  axis(side = 2, col="grey")
}
basicPlot()
```

This is first modeled with a simple linear model.

```{r icecream_glm}
lin_mod <- glm(units ~ temp, data=icecream, 
              family=gaussian(link="identity"))
library(arm) # for 'display' function only
display(lin_mod)
```

This model returns a negative intercept, which doesn't make sense. To correct this, fit a log-linear model.

```{r icecream_loglin}
log_lin_mod <- glm(log(units) ~ temp, data=icecream, 
              family=gaussian(link="identity"))
arm::display(log_lin_mod)
```

Nice - the intercept is no longer negative. The predictions from this model are reasonable to 25C, but then increase exponentially. 

```{r, echo = FALSE}
temp <- 0:35
p_log_lm <- exp(predict(log_lin_mod, data.frame(temp=0:35), type="response") + 
                  0.5 * summary(log_lin_mod)$dispersion)
p.pois <- predict(pois.mod, data.frame(temp=temp), type="response")
p.bin <- predict(bin.glm, data.frame(temp=temp), type="response")*market.size 
basicPlot(xlim=range(temp), ylim=c(0,800))
lines(temp, p_log_lm, type="l", col="red", lwd=2)
```

### Stan

Next, the blog post shows how to fit this model in a Bayesian framework using Stan.

```{r}
stanLogTransformed <-"
data {
  int N;
  vector[N] units;
  vector[N] temp;
}
transformed data {  
  vector[N] log_units;        
  log_units <- log(units);
}
parameters {
  real alpha;
  real beta;
  real tau;
}
transformed parameters {
  real sigma;
  sigma <- 1.0 / sqrt(tau);
}
model{
  // Model
  log_units ~ normal(alpha + beta * temp, sigma);
  // Priors
  alpha ~ normal(0.0, 1000.0);
  beta ~ normal(0.0, 1000.0);
  tau ~ gamma(0.001, 0.001);
}
generated quantities{
  vector[N] units_pred;
  for(i in 1:N)
    units_pred[i] <- exp(normal_rng(alpha + beta * temp[i], sigma));
}
"


temp <- c(11.9,14.2,15.2,16.4,17.2,18.1,18.5,19.4,22.1,22.6,23.4,25.1)
units <- c(185L,215L,332L,325L,408L,421L,406L,412L,522L,445L,544L,614L)
library(rstan)
stanmodel <- stan_model(model_code = stanLogTransformed)
fit <- sampling(stanmodel,
                data = list(N=length(units),
                            units=units,
                            temp=temp),
                iter = 1000, warmup=200)
stanoutput <- extract(fit)

## Extract generated posterior predictive quantities
Sims <- data.frame(stanoutput[["units_pred"]])

## Calculate summary statistics
SummarySims <- apply(Sims, 2, summary)
colnames(SummarySims) <- paste(icecream$temp,"ºC")

## Extract estimated parameters
(parms <- sapply(stanoutput[c("alpha", "beta", "sigma")], mean))

## Use parameters to predict median and mean
PredMedian <- exp(parms['alpha'] + parms['beta']*temp)
PredMean <- exp(parms['alpha'] + parms['beta']*temp + 0.5*parms['sigma']^2)

## Compare predictions based on parameters with simulation statistics
round(rbind(SummarySims, PredMedian, PredMean),1)
```

And for comparision, with RStanARM.

```{r, icecream_rstanarm}
ic_sglm <- stan_glm(log(units) ~ temp, data = icecream, family = gaussian(link = "identity"))
ic_sglm

ic_cauchy <- stan_glm(log(units) ~ temp, data = icecream, family = gaussian(link = "identity"), 
                prior = cauchy(), prior_intercept = cauchy())
ic_cauchy


posterior_interval(ic_sglm, prob = 0.95, pars = "temp")
posterior_interval(ic_cauchy, prob = 0.95, pars = "temp")
```

Well that was easy! Predictions are the same. 

